---
title: My journey in R
author: Zhichao
date: '2020-03-05'
slug: my-journey-in-r
categories: []
tags: []
---

## Exercise 1: Vectorized Code

For each part of this exercise, a function has been written for some specific purposes using loop. Your tasks are:
  - To read/run the original function to understand what it does (the easiest way to do this is to run the function through an example by hand),
  - Then write the **vectorize** the function.

- **(a)** (5 points) Name your vectorized function `specific_sum_vec()`.

```{r}
specific_sum_loop <- function(m, pos) {
  my_sum <- 0
  for (i in 1:nrow(m)) {
    for (j in pos) {
      my_sum <- my_sum + m[i, j]
    }
  }
  my_sum
}
```

Running the function:

```{r}
m <- matrix(data = c(114:133), nrow = 4, byrow = TRUE)
pos <- c(1, 4, 2)
specific_sum_loop(m, pos)
```

```{r eval = FALSE}
# write your specific_sum_vec() function
specific_sum_vec <-function(m, pos){
sum(m[,pos])
}
m <- matrix(data = c(114:133), nrow = 4, byrow = TRUE)
pos <- c(1, 4, 2)
# test your code
specific_sum_vec(m, pos)
```

- **(b)** (5 points) Name your vectorized function `long_name_vec()`.

```{r}
long_names_vec <- function(names) {
  output <- c()
  for (i in 1:length(names)) {
    name <- names[i]
    if (nchar(name) >= 8) {
      output <- c(output, name)
    }
  }
  output
}
```

Running the function:

```{r}
my_names <- c("alexander", "david", "sebastian", "johnathan", "christopher", "ha",
              "washington", "lincoln", "maximo", "mason", "luca", "anthony", "kevin")

long_names_vec(names = my_names)
```

```{r eval = FALSE}
# write your specific_sum_vec() function
long_names_vec <-function(names){
l<-nchar(names)>8
names[l]
}

# test your code
long_names_vec(names = my_names)
```


- **(c)** (10 points) Name your vectorized function `roulette_vec()`.

```{r}
roulette_loop <- function(many_bets) {
  
  # randomly decide whether you lose or win on each bet
  win_lose_random <- sample(x = c(TRUE, FALSE), size = length(many_bets),
                            replace = TRUE)
  
  total_prize <- c()
  
  # for each bet
  for (i in 1:length(many_bets)) {
    bet <- many_bets[i]
    prize <- 0
    
    # if we win, then check how much the prize is
    if (win_lose_random[i]) {
      if (bet == 'low') {
        prize <- 10
      } else if (bet == 'high'){
        prize <- 10
      } else if (bet == 'red') {
        prize <- 20
      } else if (bet == "black") {
        prize <- 20
      } else if (bet == "odd") {
        prize <- 15
      } else if (bet == "even") {
        prize <- 15
      } else if (bet == "first") {
        prize <- 50
      } else if (bet == "second") {
        prize <- 50
      } else if (bet == "third") {
        prize <- 50
      }
    }
  # add prize from this bet to the total prize
  total_prize <- c(total_prize, prize)
  }
  
  total_prize
}
```

Running the function:

```{r}
set.seed(385)
roulette_loop(many_bets = c("red"))
roulette_loop(many_bets = c("red", "black", "low", "high"))
```

```{r}
long_vec <- rep(c("red", "black", "low", "high", "second", "first", "third",
                  "odd", "even"), 10000)
system.time(roulette_loop(many_bets = long_vec))
```

- Is it taking forever to `knit` the file? If so, add `cache=TRUE` in the chunk above (that is one heavy chunk!)

```{r eval=FALSE}
# write your roulette_vec() function
roulette_vec <- function(many_bets){
   win_lose_random <- sample(x = c(TRUE, FALSE), size = length(many_bets),replace = TRUE)
   many_bets[many_bets == "low"] <- "10"
   many_bets[many_bets == "high"] <- "10"
   many_bets[many_bets == "red"] <- "20"
   many_bets[many_bets == "black"] <- "20"
   many_bets[many_bets == "even"] <- "15"
   many_bets[many_bets == "odd"] <- "15"
   many_bets[many_bets == "first"] <- "50"
   many_bets[many_bets == "second"] <- "50"
   many_bets[many_bets == "third"] <- "50"
   as.numeric(many_bets)
   many_bets<-as.numeric(many_bets) * win_lose_random
   many_bets
 }
 
# test your code
set.seed(385)
roulette_vec(many_bets = c("red"))
roulette_vec(many_bets = c("red", "black", "low", "high"))
```

Make sure to check your function using the code segment above. Then, test whether your code is faster using the code segment below:

```{r eval=FALSE}
long_vec <- rep(c("red", "black", "low", "high", "second", "first", "third",
                  "odd", "even"), 10000)
system.time(roulette_vec(many_bets = long_vec))
```

***

## Exercise 2: More Graphs!

Let's learn to graph a few more things! You might need to Google a few things, but that's a part of this course. We learn how to learn coding (with minimal guidance)!

- **(a)** (5 points) **Add a line at x = mean of `Sepal.Length`** to the following histogram

```{r fig.align='center'}
hist(x = iris$Sepal.Length, main = "Histogram of Sepal Length", xlab = "Sepal Length",
     breaks = 15, border = "dodgerblue", probability = TRUE,
     ylim = c(0, 0.65), xlim = c(4, 8.2))
box()
grid()
```

- **(b)** (5 points) Plot a **histogram of the eruption time** of the Old Faith Geyser using the `faithful` dataset, and **add a density estimate curve** to the plot.
  - *Hint*: look up `density()`
  - Use the following plot to check your answer.
  
```{r fig.align='center'}
# Plot the histogram with the density estimate curve
hist(x = iris$Sepal.Length, main = "Histogram of Sepal Length", xlab = "Sepal Length",
     breaks = 15, border = "dodgerblue", probability = TRUE,
     ylim = c(0, 0.65), xlim = c(4, 8.2))
box()
grid()
abline(v = mean(iris$Sepal.Length),col = "red",text(6.35,0.6,"x=5.843",col = "red"))
```

- **(b)** (5 points) Find a way to plot the following plots as a 1x2 matrix (shown below):
  - Histogram of waiting time from the `faithful` dataset.
  - Scatterplot of `eruptions` vs. `waiting`.
  
```{r fig.align='center'}
# plot the 2 plots
faithful
hist(x=faithful$eruption,main = "Histogram of Old Faithful Geyser Eruption Time", xlab = "Eruption Time",
     breaks = seq(from = 0, to = 6, by = 0.2), border = "dodgerblue", probability = TRUE,
     ylim = c(0, 0.8), xlim = c(1, 6))
box()
grid()
lines(density(faithful$eruption), col="blue", lwd=2)
```